Modeling Human Interactions in Collaborative Interactive Evolutionary Computation

### 2017-02-02 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-body.tex

The rest of this paper is organized as follows.
Section \ref{sec:interactive} presents related work on the topic 
of Interactive Evolution.
Then, Section \ref{sec:e} presents the Model of Human Interactions in Collaborative Interactive 
 Evolutionary Computation which 
the main proposal of this work. An implementation using the EvoSpace-Interactive  framework is detailed in Section \ref{sec:HCF}.
The case studies  and results are presented in Section \ref{sec:experiments}.
Finally, a concluding remarks are provided in Section \ref{sec:conclusions}.

\section{Related Work}
% C-IEAS
% Volunteer Based
% HBC
% Engagement Techniques
% Models

@@

\section{Human-Centered IEC Framework}
% Comment: 
% Maybe in this case we should use the term software framework or architecture. 
The general goal of this line of research is to develop a human-centered \cite{gasson2003human} 
software framework for interactive evolutionary computation (IEC). 
A framework is defined as a reusable architectural design together
with an implementation \cite{campbell1991choices}, in this case 
providing generalized components to developers of IEC systems. In the
description %below? If that's te chase, we _will_ - JJ
 we deliberately use the more general 
term Human instead of User, because humans are not always playing the role of users in software engineering sense, they
could just be interacting naturally with their environment and
unwittingly and at the same time they are implicitly
evaluating a possible solution.

@@

\section{Data Model Implementation}
The core of the framework is the data model because both the IEC system and the EA will 
depend on the application. In this section implementation details of the Data Model are presented. 
  \subsection{EvoSpace}
  \subsection{Graph}
  \subsection{PostgreSQL}

  \subsection{Interfaces}

@@

\section{Data Model Implementation}

\section{Case Studies}
\label{sec:evospace-i}
\begin{figure}[!t]
    \centering
        \includegraphics[width=3.5in]{img/interface.png}
    \caption{User interface of the EvoDrawings application.}
    \label{fig:web}
\end{figure}

@@

\subsection{Collaboration}
\label{sec:col}
Users need to authenticate themselves to the system using their
Facebook account. In that sense, some users might not be interested
either because they do not want to give that that or simply because
they do not use that particular social network. Even as we might lose
some users that way, the additional information we obtain for scoring
phenotypes more than compensates for that. % Added this, check it
                                % out. 
After entering the web application by using their Facebook account,
users can collaborate with their Facebook friends, 
sharing those phenotypes they like, or by taking phenotypes
from their friend's collections by using the web interface depicted 
in Figure \ref{fig:web}.
At the top left corner a list of Facebook friends is presented
to encourage users to interact with the system. In the central 
\emph{ Wall } area, a phenotype sampled from the population that is
being evolved via the evolutionary algorithm 
is shown to the user.
Here, the user can interact with the system in two ways.
First, he can assign a quality assessment % Why don't you use other
                                % thing instead of quality? We are not
                                % talking about quality, but about
                                % tastes or similar - JJ
 to the phenotype using
a five star rating system or,
additionally, a user can choose to add an image to one of their \emph{Collections}.
A collection is a special folder that stores those phenotypes a user likes and wishes
to save. After the user finishes interacting with the phenotype
on the Wall, he can choose to retrieve a new one from the population.
At the left hand side of \ref{fig:web}, the web page shows the \emph{Collections} section.
The user can create several collections, to group and organize his favorite 
artifacts. Moreover, users can browse the content of each collection and from
there share images through their social network.
When a user hovers the cursor over a phenotypes a pop-up pane shows how many users have
liked the phenotype. The pane also includes a link to the phenotype's 
details, the parents, genetic operators that created it, and genealogy
information. This makes the assignment of fitness through the rating
system a {\em social} activity, pursuing the objective of this work,
which is to increase engagement of the users.
\subsubsection{Graph Model} 
\subsubsection{Human Interaction Model}

@@

\section{Conclusions}

>>>>

The rest of this paper is organized as follows.
Section \ref{sec:related} presents related work on the topic 
of Interactive Evolution.
Then, Section \ref{sec:framework} presents the framework of Human Interactions in Collaborative Interactive 
 Evolutionary Computation which 
the main proposal of this work. An implementation using the EvoSpace-Interactive  framework is detailed in Section \ref{sec:implementation}.
The case studies  and results are presented in Section \ref{sec:experiments}.
Finally, a concluding remarks are provided in Section \ref{sec:conclusions}.

\section{Related Work}
\label{sec:related}
% C-IEAS
% Volunteer Based
% HBC
% Engagement Techniques
% Models

@@

\section{Human-Centered IEC Framework}
\label{sec:framework}
% Comment: 
% Maybe in this case we should use the term software framework or architecture. 
The general goal of this line of research is to develop a human-centered \cite{gasson2003human} 
software framework for interactive evolutionary computation (IEC). 
A framework is defined as a reusable architectural design together
with an implementation \cite{campbell1991choices}, in this case 
providing generalized components to developers of IEC systems. In the
description %below? If that's te chase, we _will_ - JJ
 we deliberately use the more general 
term Human instead of User, because humans are not always playing the role of users in software engineering sense, they
could just be interacting naturally with their environment and
unwittingly and at the same time they are implicitly
evaluating a possible solution.

@@

\section{Data Model Implementation}
\label{sec:implementation}

The core of the framework is the data model because both the IEC system and the EA will 
depend on the application. In this section implementation details of the Data Model are presented.
Three database models are employed to store different elements of the system:

  \subsection{EvoSpace-Redis}
EvoSpace is a population store \cite{Evospace}  for the development of 
evolutionary algorithms that are intended to run on a cloud computing model. 
The population is decoupled from any particular evolutionary algorithm. 
Candidate solutions are stored as of objects in a population, and they can be withdrawn, 
processed and replaced using a specified set of methods \cite{GValdez2015}. The population
is stored in-memory, using the Redis key-value database. Redis was chosen over a relational 
database management system, because it provides a hash based implementation of sets and 
queues which are natural data structures for the EvoSpace model. Basicaly a sample of 
candidate solutions are retieved from the server, evaluated, and then sent back. 
The same operations are used to Evolve the population. In EvoSpace individuals replaced 
in the population are stored indefinitly, to permit users to store permalinks to them.
Implementation details are presented in \cite{garcia2013evospace}.

\subsection{Graph}
  In order to deal with large datasets of connected data found in these systems,
  graph databases \cite{angles2012comparison} have been proposed as an alternative to relational databases 
  which have performance limitations when dealing with highly connected data \cite{holzschuher2013performance}.
  A graph is proposed for modeling the social network of users, their interaction with 
  candidate solutions, and the relationships between them in the population.
  The graph database system used in the implementation is Neo4J as it is
  a scalable solution \cite{miller2013graph,holzschuher2013performance}, and well supported and documented
  in PaaS infrastructures like Heroku. The Cypher query language it used to retrieve views from the graph.
  An exaplmple query is shown in figure \ref{fig:cypher} where the relations between users and solutions 
  are presented.

  \begin{figure}[!t]
    \centering
        \includegraphics[width=3.5in]{img/gui-neo.png}
    \caption{Example query in Cypher.}
    \label{fig:cypher}
  \end{figure}

\subsection{PostgreSQL}
  The PostgreSQL relational database system, is also employed because user sessions and authentication,
  as well as dynamic web pages are handled directly by the Django web framework \cite{garcia2013evospace}.

\subsection{Interfaces}
%They are Sensor and Web,

@@

\section{Case Studies}
\label{sec:experiments}

\begin{figure}[!t]
    \centering
        \includegraphics[width=3.5in]{img/interface.png}
    \caption{User interface of the EvoDrawings application.}
    \label{fig:web}
\end{figure}

@@

\subsection{Collaboration}
\label{sec:col}
Users need to authenticate themselves to the system using their
Facebook account. In that sense, some users might not be interested
either because they do not want to give that that or simply because
they do not use that particular social network. Even as we might lose
some users that way, the additional information we obtain for scoring
phenotypes more than compensates for that. % Added this, check it
                                % out. 
After entering the web application by using their Facebook account,
users can collaborate with their Facebook friends, 
sharing those phenotypes they like, or by taking phenotypes
from their friend's collections by using the web interface depicted 
in Figure \ref{fig:web}.
At the top left corner a list of Facebook friends is presented
to encourage users to interact with the system. In the central 
\emph{ Wall } area, a phenotype sampled from the population that is
being evolved via the evolutionary algorithm 
is shown to the user.
Here, the user can interact with the system in two ways.
First, he can assign a quality assessment % Why don't you use other
                                % thing instead of quality? We are not
                                % talking about quality, but about
                                % tastes or similar - JJ
 to the phenotype using
a five star rating system or,
additionally, a user can choose to add an image to one of their \emph{Collections}.
A collection is a special folder that stores those phenotypes a user likes and wishes
to save. After the user finishes interacting with the phenotype
on the Wall, he can choose to retrieve a new one from the population.
At the left hand side of \ref{fig:web}, the web page shows the \emph{Collections} section.
The user can create several collections, to group and organize his favorite 
artifacts. Moreover, users can browse the content of each collection and from
there share images through their social network.
When a user hovers the cursor over a phenotypes a pop-up pane shows how many users have
liked the phenotype. The pane also includes a link to the phenotype's 
details, the parents, genetic operators that created it, and genealogy
information. This makes the assignment of fitness through the rating
system a {\em social} activity, pursuing the objective of this work,
which is to increase engagement of the users.
\subsubsection{Graph Model}

\subsubsection{Gammification}

\subsubsection{Setup}
\subsubsection{Results}

\begin{figure}[!t]
    \centering
        \includegraphics[width=3.5in]{img/weibull_1.png}
    \caption{Experiment 1. Weibull Fit}
    \label{fig:w1}
\end{figure}
\begin{figure}[!t]
    \centering
        \includegraphics[width=3.5in]{img/weibull_2.png}
    \caption{Experiment 2. Weibull Fit}
    \label{fig:w2}
\end{figure}
\begin{figure}[!t]
    \centering
        \includegraphics[width=3.5in]{img/weibull_3.png}
    \caption{Experiment 3. Weibull Fit}
    \label{fig:w3}
\end{figure}

@@

\section{Conclusions}
\label{sec:conclusions}
